---
title: "Mapping to UKBioBank"
author: "JBF"
date: '`r Sys.time()`'
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(flextable)
library(ggplot2)
# library(dplyr)
# library(tidyr)
library(patchwork)
# library(purrr)
# library(umap)
# library(uwot)
library(gridExtra)
library(ClusterR)
library(splines)

# library(dbscan)
```

# Description
```{r}
bbdd_umap %>%
  summarise(name = 'total',
           N = as.character(n()),
           summary = '') %>%
  bind_rows(bbdd_umap %>%
              select(eid, sex, smoking) %>%
              mutate(smoking = as.character(smoking)) %>% 
              pivot_longer(cols = !eid) %>%
              group_by(name, value) %>%
              summarise(N = n(),
                        .groups = 'keep') %>%
              transmute(name,
                        value,
                        summary = sprintf('%i (%2.1f%%)', N, N/dim(bbdd_covar)[1]* 100))) %>% 
  bind_rows(bbdd_umap %>%
              select(-sex, -smoking) %>% 
              pivot_longer(cols = !eid) %>%
              group_by(name) %>%
              summarise(N = sum(!is.na(value)),
                        mn = mean(value, na.rm = T),
                        sd = sd(value, na.rm = T),
                        min = min(value, na.rm = T),
                        p025 = quantile(value, 0.025, na.rm = T),
                        q1 = quantile(value, 1/4, na.rm = T),
                        md = median(value, na.rm = T),
                        q3 = quantile(value, 3/4, na.rm = T),
                        p975 = quantile(value, 0.975, na.rm = T),
                        max = max(value, na.rm = T)) %>%
              transmute(name, 
                        N = sprintf('%i (%2.1f%%)', N, N/dim(bbdd_covar)[1]* 100),
                        summary = sprintf('%2.1f (%2.1f)', mn, sd),
                        min = sprintf('%2.1f', min),
                        p025 = sprintf('%2.1f', p025),
                        md_IQR = sprintf('%2.1f [%2.1f - %2.1f]', md, q1, q3),
                        p975 = sprintf('%2.1f', p975),
                        max = sprintf('%2.1f', max))) %>%
  select(name, value, N, summary, min, p025, md_IQR, p975, max) %>% 
  flextable %>%
  autofit
```

## Only Case-complet
```{r}
recoded_dat %>%
  summarise(name = 'total',
           N = as.character(n()),
           summary = '') %>%
  bind_rows(recoded_dat %>%
              select(eid, sex, smoking) %>%
              mutate(smoking = as.character(smoking)) %>% 
              pivot_longer(cols = !eid) %>%
              group_by(name, value) %>%
              summarise(N = n(),
                        .groups = 'keep') %>%
              group_by(name) %>%
              transmute(name,
                        value,
                        summary = sprintf('%i (%2.1f%%)', N, N/sum(N)* 100))) %>% 
  bind_rows(recoded_dat %>%
              select(-sex, -smoking) %>% 
              pivot_longer(cols = !eid) %>%
              group_by(name) %>%
              summarise(N = sum(!is.na(value)),
                        mn = mean(value, na.rm = T),
                        sd = sd(value, na.rm = T),
                        min = min(value, na.rm = T),
                        p025 = quantile(value, 0.025, na.rm = T),
                        q1 = quantile(value, 1/4, na.rm = T),
                        md = median(value, na.rm = T),
                        q3 = quantile(value, 3/4, na.rm = T),
                        p975 = quantile(value, 0.975, na.rm = T),
                        max = max(value, na.rm = T)) %>%
              transmute(name, 
                        N = sprintf('%i (%2.1f%%)', N, N/dim(recoded_dat)[1]* 100),
                        summary = sprintf('%2.1f (%2.1f)', mn, sd),
                        min = sprintf('%2.1f', min),
                        p025 = sprintf('%2.1f', p025),
                        md_IQR = sprintf('%2.1f [%2.1f - %2.1f]', md, q1, q3),
                        p975 = sprintf('%2.1f', p975),
                        max = sprintf('%2.1f', max))) %>%
  select(name, value, N, summary, min, p025, md_IQR, p975, max) %>% 
  flextable
```
## Distributions
```{r, fig.width=20, fig.height=5}
ggplot(recoded_dat %>%
         select(-smoking) %>%
         pivot_longer(-c(eid, sex)) %>%
         mutate(name = toupper(name)),
       aes(value)) +
  geom_histogram(bins = 50) +
  facet_grid(sex ~ name, scales = "free_x") +
  theme_bw()
```
## Smoking
```{r, fig.width = 5, fig.height = 5}
ggplot(recoded_dat %>%
         transmute(smoking = forcats::fct_recode(as.character(smoking),
                                                 "Yes" = "1",
                                                 "No" = "0")),
       aes(smoking)) +
  geom_bar() +
  theme_bw() +
  labs(x = "Smoking", y = "N")
```

## Corretions
```{r, fig.width=18, fig.height=9}
recoded_dat %>%
  select(-smoking) %>%
  group_by(sex) %>%
  group_modify(~{
    .x %>%
      select(-eid, -crp) %>%
      cor %>%
      reshape2::melt()
  }) %>%
  mutate(rsq = value^2, 
         to_mark = ifelse(Var1 != Var2 & rsq > .25, "*", NA),
         across(c(Var1, Var2), toupper)) %>%
  ggplot(aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = to_mark), na.rm = TRUE) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1,1)) +
  theme_minimal() +
  facet_wrap(~ sex, nrow = 1) +
  labs(x = NULL, y = NULL, fill = "Correlation", caption = "* Squared correlation higher than 0.25")
```

# Stratification by sex and transformation
## Distribution
```{r, fig.width=20, fig.height=5}
# strat_dat %>%
#     bind_rows(.id = "sex") %>%
#     pivot_longer(-c(eid, sex)) %>%
#     mutate(name = toupper(name)) %>%
#     ggplot(aes(value)) +
#     geom_histogram(bins = 50) +
#     facet_grid(sex ~ name, scales = "free_x")

ggplot(strat_predrsd,
       aes(rsd)) +
  geom_histogram(bins = 50) +
  facet_grid(sex ~ trait,
             scales = "free") +
  theme_bw()
```

### Relació entre pred i rsd
```{r, fig.width = 20, fig.height = 5}
strat_predrsd %>%
  group_by(sex) %>%
  group_map(~{
    ggplot(.x,
           aes(pred, rsd)) +
      geom_point(shape = ".", alpha = .1) +
      geom_hline(yintercept = 0, lty = "dashed", color = "red") +
      facet_wrap(~trait, nrow = 1, scales = "free") +
      theme_bw() +
      labs(x = "Predicted", y = "Residuals", title = .y)
  }) %>%
  wrap_plots(ncol = 1)
```

### QQ-plot
```{r, fig.width = 20, fig.height = 5}
ggplot(strat_predrsd,
       aes(sample = std_rsd)) +
  stat_qq(shape = ".") + 
  stat_qq_line(color = "red") +
  qqplotr::stat_qq_band() +
  facet_grid(sex ~ trait, scales = "free_x") +
  theme_bw() +
  labs(x = "Theoretical quantiles", y = "Standardised residuals")
```


## Correlations
```{r, fig.width=18, fig.height=9}
strat_dat %>%
  map_dfr(~{.x %>%
      select(-eid) %>%
      cor %>%
      reshape2::melt()
  }, .id = "sex") %>%
  mutate(rsq = value^2, 
         to_mark = ifelse(Var1 != Var2 & rsq > .25, "*", NA),
         across(c(Var1, Var2), toupper)) %>%
  ggplot(aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = to_mark), na.rm = TRUE) +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = 0, limit = c(-1,1)) +
  theme_minimal() +
  facet_wrap(~sex, nrow = 1) +
  labs(x = NULL, y = NULL, fill = "Correlation", caption = "* Squared correlation higher than 0.25")
```

## Deviation from expected relationship
```{r}
ntraits <- ncol(strat_dat[[1]]) - 1 # Subtracting the ID column
expected_mu <- rep(0, ntraits)
expected_cov <- diag(ntraits)
mahadat <- strat_dat %>%
  map(~ mahalanobis(select(.x, -eid),
                    center = expected_mu,
                    cov = expected_cov))

ggplot(mahadat %>%
         map_dfr(~ data.frame(maha_d = .x), .id = "sex"),
       aes(maha_d)) +
  geom_histogram(bins = 100, fill = "lightblue", color = "black") +
  facet_wrap(~sex, nrow = 1) +
  theme_bw() +
  labs(x = "Mahalanobis distance", y = "Count")

pvalmaha <- mahadat %>%
  map(pchisq, df = ntraits - 1, lower.tail = FALSE)

ggplot(pvalmaha %>%
         map_dfr(~data.frame(pval = .x), .id = "sex"),
       aes(pval)) +
  geom_histogram(bins = 100, fill = "lightblue", color = "black") +
  geom_vline(xintercept = .05, lty = "dashed", color = "red") +
  facet_wrap(~sex, nrow = 1) +
  theme_bw() +
  labs(x = "P-value", y = "Count")

pvalmaha %>%
  map(~round(100 * sum(.x < .05) / length(.x), 2))
```


# Map to UKBioBank
```{r}
ggplot(umap_embed %>%
         bind_rows(.id = "sex"),
       aes(X1, X2)) +
  geom_point(shape = '.') +
  facet_wrap(~sex, nrow = 1) +
  theme_bw() +
  labs(x = "UMAP1", y = "UMAP2")

ggplot(umap_embed %>%
         bind_rows(.id = "sex"),
       aes(X1, X2)) +
  geom_point(shape = '.') +
  facet_wrap(~sex, nrow = 1) +
  scale_x_continuous(limits = c(-4, 3)) +
  scale_y_continuous(limits = c(-5, 5)) +
  theme_bw() +
  labs(x = "UMAP1", y = "UMAP2")

ggplot(umap_embed %>%
         bind_rows(.id = "sex"),
       aes(X1, X2)) +
  # geom_density_2d_filled() +
  ggdensity::geom_hdr() +
  facet_wrap(~sex, nrow = 1) +
  scale_x_continuous(limits = c(-4, 3)) +
  scale_y_continuous(limits = c(-5, 5)) +
  theme_bw() +
  labs(x = "UMAP1", y = "UMAP2", alpha = "Probability")

# umap_res %>%
#     map(pluck, "embedding") %>%
#     map(data.frame) %>%
#     map2(pvalmaha, ~mutate(.x, pmaha = ifelse(.y < 0.05, " < 0.05", " > 0.05"))) %>%
#     bind_rows(.id = "sex")

# ggplot(umap_embed %>%
#          map2(pvalmaha, ~ mutate(.x, pmaha = ifelse(.y < 0.05, " < 0.05", " > 0.05"))) %>%
#          bind_rows(.id = "sex"),
#        aes(X1, X2)) +
#   geom_point(aes(color = pmaha), shape = ".") +
#   guides(color = guide_legend(override.aes = list(shape = 19, alpha = 1))) +
#   facet_wrap(~sex, nrow = 1) +
#   theme_bw() +
#     labs(x = "UMAP1", y = "UMAP2", color = "Mahalanobis\nP-value")

```

<!-- ## Outlier -->
<!-- ### Original data -->
<!-- ```{r} -->
<!-- umap_embed_fora <- umap_embed %>% -->
<!--   bind_rows(.id = "sex") %>% -->
<!--   mutate(fora = 5 < X1^2 + X2^2) -->

<!-- bbdd_umap_fora <- bbdd_umap %>% -->
<!--   inner_join(umap_embed_fora %>% select(eid, fora), -->
<!--             by = 'eid') -->

<!-- bbdd_umap_fora %>% -->
<!--   group_by(fora) %>%  -->
<!--   summarise(name = 'total', -->
<!--            N = as.character(n()), -->
<!--            summary = '') %>% -->
<!--   bind_rows(bbdd_umap_fora %>% -->
<!--               group_by(fora) %>%  -->
<!--               select(eid, fora, sex, smoking) %>% -->
<!--               mutate(smoking = as.character(smoking)) %>%  -->
<!--               pivot_longer(cols = !any_of(c('eid', 'fora'))) %>% -->
<!--               group_by(fora, name, value) %>% -->
<!--               summarise(N = n(), -->
<!--                         .groups = 'keep') %>% -->
<!--               group_by(fora, name) %>% -->
<!--               transmute(fora, -->
<!--                         name, -->
<!--                         value, -->
<!--                         summary = sprintf('%i (%2.1f%%)', N, N/sum(N)* 100))) %>%  -->
<!--   bind_rows(bbdd_umap_fora %>% -->
<!--               group_by(fora) %>%  -->
<!--               select(-sex, -smoking) %>%  -->
<!--               pivot_longer(cols = !any_of(c('eid', 'fora'))) %>% -->
<!--               group_by(fora, name) %>% -->
<!--               summarise(N = sum(!is.na(value)), -->
<!--                         mn = mean(value, na.rm = T), -->
<!--                         sd = sd(value, na.rm = T), -->
<!--                         min = min(value, na.rm = T), -->
<!--                         p025 = quantile(value, 0.025, na.rm = T), -->
<!--                         q1 = quantile(value, 1/4, na.rm = T), -->
<!--                         md = median(value, na.rm = T), -->
<!--                         q3 = quantile(value, 3/4, na.rm = T), -->
<!--                         p975 = quantile(value, 0.975, na.rm = T), -->
<!--                         max = max(value, na.rm = T)) %>% -->
<!--               transmute(fora, -->
<!--                         name,  -->
<!--                         N = sprintf('%i (%2.1f%%)', N, N/dim(bbdd_covar)[1]* 100), -->
<!--                         summary = sprintf('%2.1f (%2.1f)', mn, sd), -->
<!--                         min = sprintf('%2.1f', min), -->
<!--                         p025 = sprintf('%2.1f', p025), -->
<!--                         md_IQR = sprintf('%2.1f [%2.1f - %2.1f]', md, q1, q3), -->
<!--                         p975 = sprintf('%2.1f', p975), -->
<!--                         max = sprintf('%2.1f', max))) %>% -->
<!--   select(fora, name, value, N, summary, min, p025, md_IQR, p975, max) %>% -->
<!--   mutate(name = factor(name, -->
<!--                        levels = c('total', 'sex', 'smoking', 'age', 'alt', 'bmi', 'crp', 'dbp', -->
<!--                                   'fg', 'hdl', 'ldl', 'sbp', 'scr', 'tg', 'whr'))) %>%  -->
<!--   arrange(name, fora, value) %>%  -->
<!--   flextable %>% -->
<!--   autofit -->

<!-- bbdd_umap_fora %>% -->
<!--     pivot_longer(-c(eid, fora, sex)) %>% -->
<!--     mutate(name = toupper(name)) %>% -->
<!--     ggplot(aes(value)) + -->
<!--     geom_histogram(bins = 50) + -->
<!--     facet_grid(sex + fora ~ name, scales = "free_x") + -->
<!--     theme_bw() -->
<!-- ``` -->

<!-- ### Transformed data -->
<!-- ```{r} -->
<!-- strat_dat_fora <- strat_dat %>% -->
<!--   bind_rows(.id = "sex") %>% -->
<!--   inner_join(umap_embed_fora %>% select(eid, fora), -->
<!--             by = 'eid') -->

<!-- strat_dat_fora %>% -->
<!--   group_by(fora) %>%  -->
<!--   summarise(name = 'total', -->
<!--            N = as.character(n()), -->
<!--            summary = '') %>% -->
<!--   bind_rows(strat_dat_fora %>% -->
<!--               group_by(fora) %>%  -->
<!--               select(eid, fora, sex) %>% -->
<!--               pivot_longer(cols = !any_of(c('eid', 'fora'))) %>% -->
<!--               group_by(fora, name, value) %>% -->
<!--               summarise(N = n(), -->
<!--                         .groups = 'keep') %>% -->
<!--               group_by(fora, name) %>% -->
<!--               transmute(fora, -->
<!--                         name, -->
<!--                         value, -->
<!--                         summary = sprintf('%i (%2.1f%%)', N, N/sum(N)* 100))) %>%  -->
<!--   bind_rows(strat_dat_fora %>% -->
<!--               group_by(fora) %>%  -->
<!--               select(-sex) %>%  -->
<!--               pivot_longer(cols = !any_of(c('eid', 'fora'))) %>% -->
<!--               group_by(fora, name) %>% -->
<!--               summarise(N = sum(!is.na(value)), -->
<!--                         mn = mean(value, na.rm = T), -->
<!--                         sd = sd(value, na.rm = T), -->
<!--                         min = min(value, na.rm = T), -->
<!--                         p025 = quantile(value, 0.025, na.rm = T), -->
<!--                         q1 = quantile(value, 1/4, na.rm = T), -->
<!--                         md = median(value, na.rm = T), -->
<!--                         q3 = quantile(value, 3/4, na.rm = T), -->
<!--                         p975 = quantile(value, 0.975, na.rm = T), -->
<!--                         max = max(value, na.rm = T)) %>% -->
<!--               transmute(fora, -->
<!--                         name,  -->
<!--                         N = sprintf('%i', N), -->
<!--                         summary = sprintf('%2.1f (%2.1f)', mn, sd), -->
<!--                         min = sprintf('%2.1f', min), -->
<!--                         p025 = sprintf('%2.1f', p025), -->
<!--                         md_IQR = sprintf('%2.1f [%2.1f - %2.1f]', md, q1, q3), -->
<!--                         p975 = sprintf('%2.1f', p975), -->
<!--                         max = sprintf('%2.1f', max))) %>% -->
<!--   select(fora, name, value, N, summary, min, p025, md_IQR, p975, max) %>% -->
<!--   mutate(name = factor(name, -->
<!--                        levels = c('total', 'sex', 'smoking', 'age', 'alt', 'bmi', 'crp', 'dbp', -->
<!--                                   'fg', 'hdl', 'ldl', 'sbp', 'scr', 'tg', 'whr'))) %>%  -->
<!--   arrange(name, fora, value) %>%  -->
<!--   flextable %>% -->
<!--   autofit -->

<!-- strat_dat_fora %>% -->
<!--     pivot_longer(-c(eid, fora, sex)) %>% -->
<!--     mutate(name = toupper(name)) %>% -->
<!--     ggplot(aes(value)) + -->
<!--     geom_histogram(bins = 50) + -->
<!--     facet_grid(sex + fora ~ name, scales = "free_x") + -->
<!--     theme_bw() -->
<!-- ``` -->


## Overlaying input variables
```{r, fig.height=5, fig.width=15}
ggplot(umap_embed %>%
         map2_dfr(strat_dat,
                  inner_join,
                  by = "eid",
                  .id = "sex") %>%
         tidyr::pivot_longer(-c(X1, X2, sex, eid)) %>%
         mutate(name = toupper(name)),
       aes(X1, X2)) +
  geom_point(aes(color = value), shape = '.') +
  scale_color_gradient2(limits = c(-10, 10), low = scales::muted("blue"), high = scales::muted("red")) +
  facet_grid(sex ~ name) +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.position = "top") +
  labs(x = NULL, y = NULL, color = "Value")
```

## Overlaying age, smoking and BMI
## Age and BMI
```{r, fig.width=16, fig.height=5}
# umap_embed %>%
#   bind_rows(.id = "sex") %>%
#   {
#     d <- recoded_dat %>% 
#       select(eid, sex, age, bmi, smoking)
#     inner_join(., d, by = c("eid", "sex"))
#   } %>%
#   pivot_longer(c(age, bmi, smoking)) %>%
#   split(f = .$name) %>%
#   imap(~{
#     mp <- median(.x$value)
#     .x %>%
#       ggplot(aes(X1, X2)) +
#       geom_point(aes(color = value), shape = '.') +
#       scale_color_gradient2(midpoint = mp, high = "red", low = "blue") +
#       facet_wrap(~sex, nrow = 1) +
#       theme_bw() +
#       labs(x = "UMAP1", y = "UMAP2", 
#            title = ifelse(.y == "bmi", "BMI", stringr::str_to_title(.y)), 
#            color = NULL)
#   }) %>%
#   wrap_plots(ncol = 1)

umap_embed %>%
  bind_rows(.id = "sex") %>%
  inner_join(recoded_dat,
             by = c("eid", "sex")) %>%
  select(sex, eid, X1, X2, age, bmi) %>%
  pivot_longer(c(age, bmi)) %>%
  split(f = .$name) %>%
  imap(~{
    mp <- median(.x$value)
    ggplot(.x,
           aes(X1, X2)) +
      geom_point(aes(color = value), shape = ".") +
      scale_color_gradient2(midpoint = mp, high = "red", low = "blue") +
      facet_wrap(~sex, nrow = 1) +
      theme_bw() +
      labs(x = "UMAP1", y = "UMAP2", 
           title = ifelse(.y == "bmi", "BMI", stringr::str_to_title(.y)),
           color = NULL)
    }) %>%
    wrap_plots(nrow = 1)
```

### Smoking
```{r, fig.width=10, fig.height=5}
ggplot(umap_embed %>%
         bind_rows(.id = "sex") %>%
         inner_join(recoded_dat,
                    by = c("eid", "sex")) %>%
         transmute(sex, eid, X1, X2, 
                   smoking = forcats::fct_recode(as.character(smoking), "Yes" = "1", "No" = "0")),
       aes(X1, X2)) +
  geom_point(aes(color = smoking), shape = ".") +
  guides(color = guide_legend(override.aes = list(shape = 19))) +
  facet_wrap(~sex, nrow = 1) +
  theme_bw() +
  labs(x = "UMAP1", y = "UMAP2", color = "Smoker")
```

# Daniel's Clusters
## Males
### Subdistributions
```{r}
map2(probs,
     list(Female = data.frame(sex = "Female"),
          Male = data.frame(sex = "Male")),
     ~{
  data.frame(.x) %>%
    pivot_longer(everything()) %>%
    mutate(sex = .y$sex,
           cluster = gsub("X", "", name),
           clusterord = as.numeric(cluster))}) %>% 
  imap(~{
    ggplot(.x,
           aes(reorder(cluster, clusterord), value)) +
      geom_boxplot() +
      theme_bw() +
      labs(x = "Subdistributions",
           title = .x$sex[1])}) %>% 
    wrap_plots(nrow = 1)
```

### Maximum Probability
```{r}
wrap_plots(ggplot(maxprob$Female,
                  aes(maxprob)) +
             geom_histogram(bins = 10, fill = "lightblue", color = "black") +
             labs(title = 'Female') + 
             theme_bw(),
           ggplot(maxprob$Male,
                  aes(maxprob)) +
             geom_histogram(bins = 10, fill = "lightblue", color = "black") +
             labs(title = 'Male') + 
             theme_bw(),
           nrow = 1)
```

### Clusters in UMAP
```{r}
ggplot(umap_embed_cluster,
       aes(X1, X2)) +
  geom_point(aes(color = cluster, alpha = maxprob), shape = ".") +
  facet_wrap(~sex, nrow = 1) +
  scale_alpha_identity() +
  guides(color = guide_legend(override.aes = list(shape = 19))) +
  theme_bw() +
  labs(color = "Cluster")
```

### 95% < MaxProb
```{r}
ggplot(umap_embed_cluster %>%
         filter(0.95 < maxprob),
       aes(X1, X2)) +
  geom_point(aes(color = cluster, alpha = maxprob), shape = ".") +
  facet_wrap(~sex, nrow = 1) +
  scale_alpha_identity() +
  guides(color = guide_legend(override.aes = list(shape = 19))) +
  theme_bw() +
  labs(color = "Cluster")
```

<!-- ## Correlation between UMAP axes and input variables -->
<!-- ```{r, fig.width=10, fig.height=7} -->
<!-- umap_embed %>% -->
<!--   map2_dfr(strat_dat, inner_join, by = "eid", .id = "sex") %>% -->
<!--   { -->
<!--     d <- select(recoded_dat, eid, sex, age, smoking, bmi) -->
<!--     inner_join(., d, by = c("eid", "sex")) -->
<!--   } %>% -->
<!--   tidyr::pivot_longer(-c(sex, eid, X1, X2)) %>% -->
<!--   group_by(sex, name) %>% -->
<!--   group_modify(~{ .x %>% -->
<!--       with(data.frame(umap_axis = c("UMAP1", "UMAP2"), -->
<!--                       Correlation = c(cor(X1, value), cor(X2, value)))) -->
<!--   }) %>% -->
<!--   mutate(name = toupper(name)) %>% -->
<!--   group_by(sex, umap_axis) %>% -->
<!--   group_map(~{.x %>% -->
<!--       ggplot(aes(Correlation, reorder(name, abs(Correlation)))) + -->
<!--       geom_vline(xintercept = 0, lty = "dashed") + -->
<!--       geom_segment(aes(xend = 0, yend = reorder(name, abs(Correlation)))) + -->
<!--       geom_point(color = "red") + -->
<!--       theme_bw() + -->
<!--       labs(x = NULL, y = NULL, title = paste(.y$sex, "-", .y$umap_axis)) -->
<!--   }) %>% -->
<!--   wrap_plots -->
<!-- ``` -->

<!-- ## Archetypes -->
<!-- ```{r, fig.width=11, fig.height=5} -->
<!-- map(c("Female", "Male"), -->
<!--     ~{umap_embed[[.x]] %>% -->
<!--         ggplot(aes(X1, X2)) + -->
<!--         geom_point(shape = '.') + -->
<!--         geom_point(data = archdat_umap[[.x]], aes(fill = archnam), shape = 23, size = 3) + -->
<!--         theme_bw() + -->
<!--         labs(x = "UMAP1", y = "UMAP2", title = .x, fill = "Archetypes") -->
<!--     }) %>% -->
<!--   wrap_plots(nrow = 1) -->
<!-- ``` -->

<!-- ## Distribution of probabilities -->
<!-- ```{r, fig.width = 12, fig.height = 4} -->
<!-- arch_probs_long %>% -->
<!--     bind_rows(.id = "sex") %>% -->
<!--     ggplot(aes(archnam, prob)) + -->
<!--     geom_boxplot(aes(group = archnam, fill = archnam)) + -->
<!--     theme_bw() + -->
<!--     theme(legend.position = "none") + -->
<!--     facet_wrap(~sex, ncol = 1) + -->
<!--     labs(x = NULL, y = "Probability") -->
<!-- ``` -->

<!-- ## Maximum probabilities -->
<!-- ```{r, fig.width = 11, fig.height = 5} -->
<!-- arch_maxprob <- arch_probs_long %>% -->
<!--   map(group_by, eid) %>% -->
<!--   map(slice_max, prob, with_ties = FALSE) -->
<!-- umap_embed %>% -->
<!--     # map(pluck, "embedding") %>% -->
<!--     # map(data.frame) %>% -->
<!--     # map2(strat_dat, ~.y %>% select(eid) %>% bind_cols(.x)) %>% -->
<!--     map2_dfr(arch_maxprob, inner_join, by = "eid", .id = "sex") %>% -->
<!--     ggplot(aes(X1, X2)) + -->
<!--     geom_point(aes(color = archnam, alpha = prob), shape = '.') + -->
<!--     guides(color = guide_legend(override.aes = list(shape = 19))) + -->
<!--     scale_alpha_identity() + -->
<!--     facet_wrap(~sex, nrow = 1) + -->
<!--     theme_bw() + -->
<!--     labs(x = "UMAP1", y = "UMAP2", color = "Archetypes") -->
<!-- ``` -->

<!-- ```{r, fig.width = 20, fig.height = 5} -->
<!-- arch_maxprob %>% -->
<!--   imap(~{ -->
<!--     d <- umap_embed[[.y]] -->
<!--     .x %>% -->
<!--       split(f = .$archnam) %>% -->
<!--       map(function(ARCH){ -->
<!--         dd <- inner_join(ARCH, d, by = "eid") -->
<!--         d %>% -->
<!--           ggplot(aes(X1, X2)) + -->
<!--           geom_point(shape = '.', alpha = .1) + -->
<!--           geom_point(data = dd, color = "red", alpha = .5, size = .2) + -->
<!--           theme_bw() + -->
<!--           labs(x = NULL, y = NULL, title = ARCH$archnam[1]) -->
<!--         }) %>% -->
<!--       modify_at(1, function(g){ g + labs(y = .y) }) %>% -->
<!--       wrap_plots(nrow = 1) -->
<!--     }) %>% -->
<!--   wrap_plots(nrow = 2) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- umap_embed %>% -->
<!--   map2_dfr(arch_maxprob, inner_join, by = "eid", .id = "sex") %>% -->
<!--   filter(prob > .5) %>% -->
<!--   ggplot(aes(X1, X2)) + -->
<!--   geom_point(aes(color = archnam, alpha = prob), size = .25) + -->
<!--   guides(color = guide_legend(override.aes = list(size = 1))) + -->
<!--   scale_alpha_identity() + -->
<!--   facet_wrap(~sex, nrow = 1) + -->
<!--   theme_bw() + -->
<!--   labs(x = "UMAP1", y = "UMAP2", color = "Archetypes") -->
<!-- ``` -->

<!-- ## Overlaying prevalent diseases -->
<!-- ```{r, fig.width = 20, fig.height = 10} -->
<!-- umap_disease %>% -->
<!--     imap(~{ -->
<!--       .x %>% -->
<!--         tidyr::pivot_longer(any_of(c("angor", "tia", "stroke", "ami", "COPD", "CKD", "cancer", -->
<!--                                      "depress", "htn", "hf", "liver", "ra", "sleep_apnea", -->
<!--                                      "pcos")), -->
<!--                             names_to = "Disease", -->
<!--                             values_to = "Diagnosis") %>% -->
<!--         mutate(Disease = gsub("_", " ", Disease), -->
<!--                Diagnosis = ifelse(Diagnosis == 1, "Yes", "No")) %>% -->
<!--         ggplot(aes(X1, X2)) + -->
<!--         geom_point(aes(color = Diagnosis,  -->
<!--                        shape = ifelse(Diagnosis == "Yes", "*", "."), -->
<!--                        alpha = ifelse(Diagnosis == "Yes", .5, .1)), -->
<!--                    na.rm = TRUE) + -->
<!--         scale_shape_identity() + -->
<!--         scale_alpha_identity() + -->
<!--         scale_color_manual(values = c("grey", "red")) + -->
<!--         guides(color = guide_legend(override.aes = list(alpha = 1))) + -->
<!--         geom_point(data = archdat_umap[[.y]], aes(fill = archnam), shape = 23, size = 3) + -->
<!--         facet_wrap(~Disease, nrow = 2) + -->
<!--         theme_bw() + -->
<!--         labs(title = .x$sex[1], x = "UMAP1", y = "UMAP2", fill = paste("Archetypes -", .y)) -->
<!--     }) %>% -->
<!--   patchwork::wrap_plots(ncol = 1, guides = "collect") -->
<!-- ``` -->

